The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-11-28 article_final/main.tex
Minor improvements

\title[Ky Fan Norms for Matrix Optimization]{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}

@@

We are particularly interested in the case when $\cS$ is a unit ball in some norm $\norm{\cdot}$:
\[
\cS = \cB_{\norm{\cdot}} = \{ \mD \in \Rmn \mid \norm{\mD} \leq 1 \}\,.
\]
In this case,
\[
\argmin_{\mD \in \cS} \<\mM^t, \mD> = - \{ \mD \in \cB_{\norm{\cdot}} \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,,
\]
and the update for $\mX^t$ in \equationref{eq:simple_update} simplifies to
\begin{equation}\label{eq:our_update}
	\mX^{t+1} = \mX^{t} - \gamma_t \{\mD \in \cB_1 \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,.
\end{equation}

@@

Let us derive the update for an arbitrary $k$. Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p. 96), the goal is to reach $\max\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i, \sigma_1\}$. $\Delta = \vu_1 \vv_1^\top$ from Neon delivers $\sigma_1$, while $\Delta = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ from Muon (with an extra factor of $1/k$) delivers $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$. Thus, the update is either
\[
\mX^{t+1} = \mX^{t} - \gamma_t \vu_1 \vv_1^\top\text{ or }\mX^{t+1} = \mX^{t} - \frac{\gamma_t}{k}\mU \mV^\top\,,
\]
depending on which one better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \mM^t(\mX - \mX^t)$.

>>>>

\title[Ky Fan Norms, Duals, and Norm Combinations for Matrix Optimization]{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}

@@

We are particularly interested in the case when $\cS$ is a unit ball in some norm $\norm{\cdot}$:
\[
\cS = \cB_{\norm{\cdot}} = \{ \mD \in \Rmn \mid \norm{\mD} \leq 1 \}\,.
\]
In this case,
\[
\argmin_{\mD \in \cS} \<\mM^t, \mD> = - \{ \mD \in \cB_{\norm{\cdot}} \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,,
\]
and the update for $\mX^t$ in \equationref{eq:simple_update} simplifies to
\begin{equation}\label{eq:our_update}
	\mX^{t+1} = \mX^{t} - \gamma_t \{\mD \in \cB_{\norm{\cdot}} \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,.
\end{equation}

@@

Let us derive the update for an arbitrary $k$. Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p. 96), the goal is to reach $\max\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i, \sigma_1\}$. $\Delta = \vu_1 \vv_1^\top$ from Neon delivers $\sigma_1$, while $\Delta = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ from Muon (with an extra factor of $1/k$) delivers $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$. Thus, the update is either
\[
\mX^{t+1} = \mX^{t} - \gamma_t \vu_1 \vv_1^\top\text{ or }\mX^{t+1} = \mX^{t} - \frac{\gamma_t}{k}\mU \mV^\top\,,
\]
depending on which one better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \<\mM^t, \mX - \mX^t>$.
